\documentclass[12pt, twoside]{report}
\usepackage[utf8]{inputenc}
\usepackage[a4paper, top=30mm, left=20mm, bottom=20mm,
    right=20mm]{geometry}
\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage{color}
\usepackage{fancyhdr}
\fancyhead[LO,RE]{\itshape \nouppercase Chapter \arabic{chapter}}
\usepackage{amssymb}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{mathtools} % shortintertext
\usepackage{faktor}
\usepackage[backend=bibtex, 
            maxbibnames=10,
            style=alphabetic]{biblatex}
%\renewcommand{\familydefault}{\sfdefault}
%\usepackage{helvet}
\usepackage{hyperref}
\usepackage{subfigure}
\usepackage[font={small,it}]{caption}

\usepackage{algpseudocode}
\usepackage[plain]{algorithm}
\usepackage[shortlabels]{enumitem}
\usepackage{booktabs}
\usepackage{mhchem} % chemical reactions
\usepackage{cleveref} % multireference (eq 1.2-1.5)
\usepackage{pgfgantt} % nice Gantt diagrams
\usepackage{bm}
% teste

\pagestyle{fancy}

% Images path
\graphicspath{ {img/} }

% ABNT foreign words should be in italic
\newcommand{\foreignword}[1]{\textit{#1}}
\newcommand{\toolname}[1]{\textit{#1}}
\newcommand{\fieldR}{\mathbb{R}}
\newcommand{\naturals}{\mathbb{N}}
\newcommand{\powerset}{\mathcal{P}}
\newcommand{\probability}{\mathbb{P}}
\newcommand{\expectation}{\mathbb{E}}
\newcommand{\algname}[1]{\texttt{#1}}
\newcommand{\langname}[1]{\texttt{#1}}
\newcommand{\varname}[1]{\texttt{#1}}
\newcommand{\floor}[1]{\lfloor #1 \rfloor}
\newcommand{\ceil}[1]{\lceil #1 \rceil}
\newcommand{\mathsc}[1]{{\normalfont\textsc{#1}}}
\newcommand{\forest}{\mathcal{F}}
\newcommand{\pfsnode}[1]{\mathbf #1}
\newcommand{\species}[1]{\textit{#1}}
\newcommand{\gender}[1]{\textit{#1}}

% remove returns of the same line in pseudocodes
\algrenewcommand\Return{\State \algorithmicreturn{} } 

\DeclareMathOperator*{\argmin}{argmin} 

\addbibresource{references.bib}

\newtheorem{mydefinition}{Definition}
\numberwithin{mydefinition}{section}
\newtheorem{mytheorem}{Theorem}
\numberwithin{mytheorem}{section}
\newtheorem{mylemma}{Lemma}
\numberwithin{mylemma}{section}
\newtheorem{corollary}{Corollary}
\numberwithin{corollary}{section}


\begin{document}
\pagenumbering{roman}

\thispagestyle{empty}
\begin{center}
{\Large
{\bf Identification of cell signaling pathways based on biochemical 
    reaction kinetics repositories}\\
\bigskip
\bigskip
\bigskip
\bigskip
    {\bf \href{mailto:gustavo.estrela.matos@gmail.com}{Gustavo Estrela de Matos}}\\
\bigskip
\bigskip
\bigskip
\bigskip
\textsc{
    Text presented\\[-0.25cm] 
    to\\[-0.25cm]
    Institute of Mathematics and Statistics\\[-0.25cm]
    of the\\[-0.25cm]
    University of São Paulo\\[-0.25cm]
    For\\[-0.25cm]
    Obtaining the Title Of Master Of Science\\
    }
\bigskip
\bigskip
\bigskip
\bigskip
Field of knowledge: Computer Science\\
\bigskip
Advisor: Dr. Marcelo da Silva Reis\\
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
Center of Toxins, Immune-Response and Cell Signaling (CeTICS)\\
\bigskip
Special Laboratory of Cell Cycle, Butantan Institute\\
\bigskip
\bigskip
{\normalsize During the development of this work the author received 
    financial support from scholarship \#~17/20575-9, São Paulo Research Foundation (FAPESP).}\\
\bigskip
\bigskip
\bigskip
São Paulo, \today
}
\end{center}
\newpage

\thispagestyle{empty}
\begin{center}
{\Large
{\bf Identificação de vias de sinalização celular baseada em
    repositórios de cinética de reações bioquímicas}\\
\bigskip
\bigskip
\bigskip
\bigskip
    {\bf \href{mailto:gustavo.estrela.matos@gmail.com}{Gustavo Estrela de Matos}}\\
\bigskip
\bigskip
\bigskip
\bigskip
\textsc{
    Texto apresentado\\[-0.25cm] 
    ao\\[-0.25cm]
    Instituto de Matemática e Estatística\\[-0.25cm]
    da\\[-0.25cm]
    Universidade de São Paulo\\[-0.25cm]
    para\\[-0.25cm]
    Obtenção do Título de\\
    Mestre em Ciências\\
    }
\bigskip
\bigskip
\bigskip
\bigskip
Área do Conhecimento: Ciência da Computação\\
\bigskip
Orientador: Dr. Marcelo da Silva Reis\\
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
Center of Toxins, Immune-Response and Cell Signaling (CeTICS)\\
\bigskip
Special Laboratory of Cell Cycle, Butantan Institute\\
\bigskip
\bigskip
{\normalsize Durante o desenvolvimento deste projeto, o autor recebeu
    suporte financeiro através da agência FAPESP (processo 17/20575-9).}\\
\bigskip
\bigskip
\bigskip
São Paulo, 12 de janeiro de 2021
}
\end{center}
\newpage


\thispagestyle{empty}
\begin{center}
{\Large
{\bf Identificação de vias de sinalização celular baseada em
    repositórios de cinética de reações bioquímicas}\\
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
    {\href{mailto:gustavo.estrela.matos@gmail.com}{Gustavo Estrela de Matos}}\\
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\hfill
\begin{minipage}{15em}
{\small Esta é a versão original da dissertação\\
elaborada pelo candidato Gustavo Estrela de Matos,
tal como submetida à Comissão Julgadora.}
\end{minipage}
\bigskip
\bigskip
}
\end{center}
\newpage

\chapter*{Abstract}
Cell signaling pathways are composed of a set of biochemical reactions 
that are associated with signal transmission within the cell and its 
surroundings. Traditionally, those pathways are identified through 
statistical analyses on results from biological assays, in which 
involved chemical species are quantified. However, once generally it is 
measured only a few time points for a fraction of the chemical species, 
to effectively tackle this problem it is required to design and simulate 
functional dynamic models. Recently, a method was introduced to 
design functional models, which is based on systematic modifications of 
an initial model through the inclusion of biochemical reactions, which 
in turn were obtained from the interactome repository KEGG. 
Nevertheless, this method presents some shortcomings that impair the 
estimated model; among them are the incompleteness of the information 
extracted from KEGG, the absence of rate constants, the usage of 
sub-optimal search algorithms and an unsatisfactory overfitting 
penalization. In this work, we propose a new methodology for 
identification of cell signaling pathways, based on the aforementioned
method, with modifications on the cost function that aims to solve the
unsatisfactory overfitting penalization. To this end, we defined a cost function
based on the marginal likelihood of a model producing the observed
experimental data. We show how this new cost function 
automatically penalize complex models, since marginal likelihood 
approaches tend to select models with intermediate complexity. The new 
methodology was tested on artificial instances of model selection; for 
one of the experiments, we solved the model selection problem as a 
feature selection problem, walking on the space of solutions to get a 
glance of the surface induced by the defined cost function. With this
work, we expect to contribute towards the solution of the cell 
signaling pathway identification problem, by providing the implementation of a cost function that
can be used in a feature selection approach.

\chapter*{Resumo}
Vias de sinalização celular são compostas por um conjunto de reações
bioquímicas que estão associadas à transmissão de informação no interior
de uma célula e suas imediações. Tradicionalmente, essas vias são
identificadas com análises estatísticas sobre resultados de ensaios
biológicos que quantificam espécies químicas envolvidas. Todavia, como
geralmente são medidos apenas alguns instantes de tempo de uma fração
dessas espécies químicas, para efetivamente abordar esse problema é
necessário o desenho e a simulação de modelos dinâmicos funcionais.
Recentemente, foi introduzido um método para desenho de modelos
funcionais baseado em modificações sistemáticas de um modelo inicial
através da inclusão de reações bioquímicas extraídas do repositório de
interatomas KEGG. Entretanto, este método apresenta limitações que
comprometem o modelo estimado; entre elas, a incompletude das
informações extraídas do KEGG, a ausência de constantes de velocidade, o
uso de algoritmos de busca subótimos e uma penalização insatisfatória
para sobre-ajuste. Neste trabalho, propomos uma nova metodologia para
identificação de vias de sinalização celular, baseada no método citado,
com modificações na função de custo que tem como objetivo penalizar
complexidade de modelos satisfatoriamente. Para isso, definimos uma
função de custo baseada na verossimilhança marginal de um modelo 
reproduzir dados experimentais. Produzimos experimentos e mostramos como
a nova função de custo automaticamente penaliza modelos mais complexos,
o que é esperado pois abordagens baseadas em verossimilhança marginal
tendem a selecionar modelos de complexidade intermediária. Nossa
metodologia foi testada em instancias artificiais de seleção de modelos;
em um dos experimentos, realizamos a seleção de modelos como um problema
de seleção de características, caminhando pelo espaço de soluções para
se obter indícios sobre a superfície induzida pela função de custo sobre
o espaço de busca. Com este trabalho, esperamos contribuir para a 
elucidação do problema de seleção modelos de vias de sinalização celular,
fornecendo uma função de custo que possa ser utilizada em uma abordagem
baseada em seleção de características.

\tableofcontents

\clearpage
\pagenumbering{arabic} 

\nocite{*}
\chapter{Introduction}
\label{chap:intro}
\input{chapters/introduction}

\chapter{Fundamental concepts}
\label{chap:fundamental_concepts}
\input{chapters/fundamental_concepts}

\chapter{Model selection methods}
\label{chap:model_selection}
\input{chapters/model_selection}

\chapter{Development of SigNetMS, a software for model selection}
\label{chap:development_signetms}
\input{chapters/development_signetms}

\chapter{Experiments and results}
\label{chap:experiments}
\input{chapters/experiments}

\chapter{Conclusion}
\label{chap:conclusion}
\input{chapters/conclusion}

\newpage
\printbibliography

\end{document}

